---
title: "Statistical Inference Project"
author: Sloane Simmons
date: November 23, 2014
output: pdf_document
---
Statistical Inference Project - Exponential Distributions
========================================================

We can generate a vector of observations with the R function `rexp(n,lambda)`:

```{r,results='markup',eval=TRUE,echo=TRUE}
rexp(40,0.2)
```

We generate a large number of these vectors (arbitrarily chosen to be 1000) and then take the mean of each:
```{r,results='markup',eval=TRUE,echo=TRUE}
expmeans <- 1:1000

for (i in 1:1000) { 
  expmeans[i] <- mean(rexp(40,0.2))
}

head(expmeans)
```

Exponential distributions are centered at $1/\lambda$ with variance $1/(\lambda^2)$; for $\lambda$ = 0.2, this gives a $\mu = 5$ with $\sigma^2 = 25$ ($\sigma = 5.0$).  From the Law of Large numbers, our distribution of sample means should be centered at $\mu$ with a standard deviation of $\frac{\sigma}{\sqrt{n}}$, which here is: 
```{r,eval=TRUE} 
sqrt(5/8)
```

Comparing this to our results, we see it is centered at `mean(expmeans)`=`r mean(expmeans)` with standard deviation `sd(expmeans)`=`r sd(expmeans)`.  This compares well with the theoretical distribution (within one standard deviation).

```{r}
sdist <- sd(expmeans)
samplemean <- mean(expmeans)
(samplemean - 5.0)/(5/sqrt(40)) # Number of standard deviations sample mean is from calculated mean
```

We can see fairly easily from a summary and histogram of the data that it is roughly fits a normal distribution centered at 5.  (Even though each indvidual measurement is of an exponential variable, the means of these measurements (due to the law of large numbers) follows a normal distribution, with sample mean centered near that of the population mean.)

```{r fig.width=7, fig.height=6}
summary(expmeans)
hist(expmeans)
```

