---
title: "Statistical Inference Project"
author: Sloane Simmons
date: November 23, 2014
output: pdf_document
---
Statistical Inference Part 2: 
========================================================
```{r,echo=FALSE,results='hide'}
suppressMessages(library(dplyr))
suppressMessages(library(tidyr))
```

### Analysis

We begin by loading the ToothGrowth data:
```{r}
tg <- ToothGrowth
summary(tg)
```

```{r,echo=FALSE,results='hide'}
tg <- tbl_df(tg)
```
It records (in the 'len' column), the growth experienced by the Guinea Pig with a certain treatment (supplement type and dose).  We can use the t test to compare tooth growth by supplement type and by dosage, but first we'll need to separate out the data using dplyr:

```{r,echo=TRUE}
# First separate out observations by supplement type
OJ <- as.vector(tg %>% filter(supp=="OJ") %>% dplyr::select(len))$len
VC <- as.vector(tg %>% filter(supp=="VC") %>% dplyr::select(len))$len

# Observations by dosage
tg$dose <- as.factor(tg$dose)
# Will be 3-elem list with 0.5,1,2 from factors of dosages
# data we want is in doses[[i]]$len where i is 1 to 3
doses <- split(tg,tg$dose)

```
#### By Supplement type:
Our null hypothesis $H_0$ is that the means of the length for each supplement type is equal.  The alternative hypothesis $H_\alpha$ is that they are not.  We will reject with $\alpha = 0.05$ using an two-sided, unpaired t-test (*not* assuming that the variances are equal.)
```{r}
t.test(OJ,VC,alternative="two.sided",var.equal=FALSE)

```

Conclusion:
Because this gives a p-value of ~`0.06`>`0.05`, we **cannot** reject $H_0$, so we conclude that the supplement type makes no difference, and that the means are equal.

#### By dosage:
This time, we have 3 samples to compare (one for each dosage).  We will use the same assumptions (unpaired, two-sided, $\alpha = 0.05$) in comparing them:
```{r}
# doses 1:3 is 0.5,1,2 levels
t.test(doses[[1]]$len,doses[[2]]$len,alternative="two.sided",var.equal=FALSE)
t.test(doses[[1]]$len,doses[[3]]$len,alternative="two.sided",var.equal=FALSE)
t.test(doses[[2]]$len,doses[[3]]$len,alternative="two.sided",var.equal=FALSE)
```

Conclusion:
The p-values for all of these are $<< 0.05$, so we *reject* the null hypotheses and conclude that the means are different (dosage *is* significant).

### Summary
We can best see the length differences by using a boxplot:

```{r}
boxplot(len~dose, data= ToothGrowth)
```

